Finch: Sparse and Structured Array Programming with Control Flow

### 2024-03-25 sections/lang.tex
adding captions, tying things together.

\noindent % This ensures the minipages fill the page width
\begin{minipage}{0.4\linewidth}
\begin{align*}
    \finchliteral(val \in \mathbb{V}) &:= \text{\mintinline{julia}{val}} \\
    \finchvalue(ex \in \mathbb{S}, type \in \mathbb{T}) &:= \text{\mintinline{julia}{ex :: type}} \\
    \finchtensor(name \in \mathbb{S}) &:= \text{\mintinline{julia}{name}} \\
    \finchindex(name \in \mathbb{S}) &:= \text{\mintinline{julia}{name}} \\
    \finchvar(name \in \mathbb{S}) &:= \text{\mintinline{julia}{name}} \\
    \finchextent(a \in E, b \in E) &:= \text{\mintinline{julia}{a : b}} \\
    \finchcall(f \in E, args\ldots \in E) &:= \text{\mintinline{julia}{f(args...)}} \\
    \finchaccess(tns \in T, idxs\ldots \in E) &:= \text{\mintinline{julia}{tns[idxs...]}} \\
    \finchfreeze(tns \in T) &:= \text{\mintinline{julia}{@freeze(tns)}} \\
    \finchthaw(tns \in T) &:= \text{\mintinline{julia}{@thaw(tns)}} \\
    \finchmode(tns \in T) &:= \text{\mintinline{julia}{@mode(tns)}} \\
\end{align*}
\end{minipage}%
\begin{minipage}{0.5\linewidth}
\begin{align*}
    \mathbb{V} &:= \text{the set of all values.} \\
    \mathbb{S} &:= \text{the set of all symbols.} \\
    \mathbb{T} &:= \text{the set of all types.} \\
    I &:= \finchindex \\
    A &:= \finchaccess \\
    V &:= \finchvar \\
    T &:= \finchtensor \\
    E &:= \finchliteral \mid \finchvalue \mid \finchindex \\
        &\quad \mid \finchvar \mid \finchextent \mid \finchcall \mid \finchaccess \\
    S &:= \finchassign \mid \finchloop \mid \finchdefine \\
        &\quad \mid \finchsieve \mid \finchblock \\
\end{align*}
\end{minipage}%
\begin{align*}
    \finchdeclare(tns \in T, init \in E, dims\ldots \in E) &:= \text{\mintinline{julia}{tns .= init(dims...)}} \\
    \finchassign(lhs \in A, op \in E, rhs \in E) &:= \text{\mintinline{julia}{lhs \texttt{<<op>>=} rhs}} \\
    \finchloop(idx \in I, range \in E, body \in S) &:= \text{\mintinline{julia}{for idx = range; body end}} \\
    \finchdefine(var \in V, val \in E, body \in S) &:= \text{\mintinline{julia}{let var = val; body end}} \\
    \finchsieve(cond \in E, body \in S) &:= \text{\mintinline{julia}{if cond; body end}} \\
    \finchblock(bodies\ldots \in S) &:= \text{\mintinline{julia}{begin; bodies... end}}
\end{align*}

>>>>

The syntax of Finch is displayed in Figure \ref{fig:syntax}. We

\begin{figure}
\noindent % This ensures the minipages fill the page width
\begin{minipage}{0.4\linewidth}
\begin{align*}
    \finchliteral(val \in \mathbb{V}) &:= \text{\mintinline{julia}{val}} \\
    \finchvalue(ex \in \mathbb{S}, type \in \mathbb{T}) &:= \text{\mintinline{julia}{ex :: type}} \\
    \finchtensor(name \in \mathbb{S}) &:= \text{\mintinline{julia}{name}} \\
    \finchindex(name \in \mathbb{S}) &:= \text{\mintinline{julia}{name}} \\
    \finchvar(name \in \mathbb{S}) &:= \text{\mintinline{julia}{name}} \\
    \finchextent(a \in E, b \in E) &:= \text{\mintinline{julia}{a : b}} \\
    \finchcall(f \in E, args\ldots \in E) &:= \text{\mintinline{julia}{f(args...)}} \\
    \finchaccess(tns \in T, idxs\ldots \in E) &:= \text{\mintinline{julia}{tns[idxs...]}} \\
    \finchfreeze(tns \in T) &:= \text{\mintinline{julia}{@freeze(tns)}} \\
    \finchthaw(tns \in T) &:= \text{\mintinline{julia}{@thaw(tns)}} \\
    \finchmode(tns \in T) &:= \text{\mintinline{julia}{@mode(tns)}} \\
\end{align*}
\end{minipage}%
\begin{minipage}{0.5\linewidth}
\begin{align*}
    \mathbb{V} &:= \text{the set of all values.} \\
    \mathbb{S} &:= \text{the set of all symbols.} \\
    \mathbb{T} &:= \text{the set of all types.} \\
    I &:= \finchindex \\
    A &:= \finchaccess \\
    V &:= \finchvar \\
    T &:= \finchtensor \\
    E &:= \finchliteral \mid \finchvalue \mid \finchindex \\
        &\quad \mid \finchvar \mid \finchextent \mid \finchcall \mid \finchaccess \\
    S &:= \finchassign \mid \finchloop \mid \finchdefine \\
        &\quad \mid \finchsieve \mid \finchblock \\
\end{align*}
\end{minipage}%
\begin{align*}
    \finchdeclare(tns \in T, init \in E, dims\ldots \in E) &:= \text{\mintinline{julia}{tns .= init(dims...)}} \\
    \finchassign(lhs \in A, op \in E, rhs \in E) &:= \text{\mintinline{julia}{lhs \texttt{<<op>>=} rhs}} \\
    \finchloop(idx \in I, range \in E, body \in S) &:= \text{\mintinline{julia}{for idx = range; body end}} \\
    \finchdefine(var \in V, val \in E, body \in S) &:= \text{\mintinline{julia}{let var = val; body end}} \\
    \finchsieve(cond \in E, body \in S) &:= \text{\mintinline{julia}{if cond; body end}} \\
    \finchblock(bodies\ldots \in S) &:= \text{\mintinline{julia}{begin; bodies... end}}
\end{align*}
\caption{The syntax of the finch language. Note that instead of a classic EBNF
form, we have constrained the classes of subterms on the left hand side to allow
for descriptive names of the arguments. Compare this grammar to the Concrete
Index Notation of TACO \cite[Figure~3]{kjolstad_tensor_2019}, noting the
addition of multiple left hand sides through code blocks, and the explicit
initialization and finalization of tensors.}\label{fig:syntax}
\end{figure}
